On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-02-19 MS.tex
Address comments from Mark Westoby and Freya Thomas

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoretical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 others traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenetic stages: seedling, sapling and adult.

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across ontogenetic stages. Specific leaf area was correlated with relative growth rate in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with relative growth rate only in one ontogenetic stage category: seedlings and adults, respectively.

\item Although we were able to confirm several of our theoretical predictions, major knowledge gaps still exist in the trait literature. For example, for one-third of the trait considered, the majority ($>$ 75\%) of reported correlations with growth came from a single ontogenetic stage.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that a strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate  among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that the trait specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. From the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents, latitudes, and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of trait variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, recently a number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an example. By definition, species with high SLA can deploy more leaf area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In seedlings growing under favourable conditions, SLA has repeatedly been found to be a strong predictor of relative growth rate, regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{King-1999, Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organism design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta, King-2011}. As plants grow larger, a greater amount of the energy is devoted towards building and maintaining non-photosynthetic tissues \citep{Givnish:1995ta,King-2011}. A recent model by \citet{Falster:2011ii} suggests this change shifts the benefits and costs of some trait-based trade-offs, such that the net effect of the trait on growth changes with size. Thus the absence of a correlation between SLA and growth at larger sizes may in fact be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. While recent empirical studies support the idea that the relationship between traits and growth rates may change with size (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood (or stem) tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum (or asymptotic height), and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

Below we first outline equations for the new growth model, adapted from \citet{Falster:2011ii}. We then use this model to generate qualitative predictions on how the effect of traits on growth rate changes with plant size. These hypotheses are summarised in Table \ref{tab:trade-offs}.

@@

The rate at which \textbf{live biomass} $B$ is produced by a plant is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}} \, ,
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i = \rm{l,s,r,t,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes \textbf{areas}, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm t} + M_{\rm f}$.  Applying the chain rule of differential calculus, the \textbf{growth rate in plant mass}, ($M_{\rm t}$), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm t}}{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm t}}{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm t}}{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The \textbf{growth rate in height} can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm t}}}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm t}}{\ud B}} \times {\frac{\ud B}{\ud t}}.
\end{equation}
where $\ud H / \ud A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. We assume this function is unrelated to the traits considered in the current paper. The second term, $\ud A_{\rm l} / \ud M_{\rm t}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\ud A_{\rm l} / \ud M_{\rm t}$
can itself be expressed as a sum of construction costs per unit leaf
area,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm t}}
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

Similarly \textbf{growth rate in basal area} can be expressed as the sum of growth in sapwood and heartwood, with the former term again decomposed as a product:
\begin{equation}\label{eq:eq3}
\begin{array}{ccccc}
\underbrace{{\frac{\ud A_{\rm st}}{\ud t}}}_{\text{basal area growth rate}} &=& \frac{\ud A_{\rm ss}}{\ud t} &+& \frac{\ud A_{\rm sh}}{\ud t} \\
&=& \underbrace{\frac{\ud A_{\rm ss}}{\ud A_{\rm l}}  \times {\frac{\ud A_{\rm l}}{\ud M_{\rm t}}} \times  {\frac{\ud M_{\rm t}}{\ud B}} \times {\frac{\ud B}{\ud t}}}_{\text{increase in total sapwood}} &+&  \underbrace{k_{\rm ss} \times \frac{\ud A_{\rm ss}}{\ud t}}_{\text{sapwood turnover}},\\
\end{array}
\end{equation}
where $\frac{\ud A_{\rm ss}}{\ud A_{\rm l}} $ is the  sapwood area per leaf area. The \textbf{growth rate in diameter} is then obtained via the geometric relationship between stem diameter and stem area,
\begin{equation}\label{eq:eq4}
\underbrace{\frac{\ud D}{\ud t}}_{\text{growth rate in diameter}} = \sqrt[]{\frac{\Pi}{A_{\rm st}}} \times {\frac{\ud A_{\rm st}}{\ud t}}.
\end{equation}

@@

The above equations are general, i.e. they do not depend on detailed assumptions about plant construction. To make qualitative predictions with respect to traits such as specific leaf area and wood density, however, we require one additional assumption: that the mass of sapwood tissue ($M_{\rm ss}$) increases disproportionately with respect to the plants leaf area ($A_{\rm l}$). This assumption, which is included in nearly every model of plant growth that has been developed, has the effect of changing the relative importance of different components within equations for $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ above as a plant increases in size. Specifically, in eq. \ref{eq:eq0}, the costs of stem respiration and turnover increase faster than leaf photosynthesis, respiration or turnover, as the plant grows. Similarly, in eq. \ref{eq:daldmt}, the costs of constructing sapwood support to support a given unit of leaf area, $\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$, increases faster the cost of producing a unit of leaf area, $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ which is simply the inverse of specific leaf area. Whether plant growth rate increases with a specific trait, therefore depends on how the trait effects the different elements of \ref{eq:eq0}-\ref{eq:eq4} and how those effects change with size (Table \ref{tab:trade-offs}).

Based on the above model and this single assumption about growth, we outline five hypotheses that arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} and how these may interact with plant size (see Table \ref{tab:trade-offs}). These hypotheses reflect a direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. Nevertheless, for purposes of this paper we adopt as working hypotheses only those arising from direct mechanistic causation. Moreover, we consider how predictions may vary when expressed on absolute and relative terms, for each of height, diameter, and mass-based growth (see Table \ref{tab:trade-offs}). For many trait and growth combinations, the predicted effect of the trait on growth is similar when expressed on relative and absolute terms. However, this is not true for all traits (details given below). We therefore adopted a two-stage approach to our analysis, whereby we first assess patterns based on measurements of relative growth rate.

\subsubsection*{H1: The relationship between SLA and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants}

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA change with size (Table \ref{tab:trade-offs}). 
During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is negative. This occurs because of the increased turnover of high SLA species.

@@

Cheaper stem construction (low wood density or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is non-significant along all plant size\citep[see also][]{Enquist-1999}.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq4}). For seedlings, where all individuals are investing solely in growth, we expect no influence of maximum height on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their maximum height value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller maximum height are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of  growth is used (Table \ref{tab:trade-offs}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have larger absolute growth rates. However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based growth rates during the earliest phases of growth \citep{Turnbull:2012ew}. Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therfore disappear. Thus for larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

\subsection*{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is typically measured. This effects occurs within the term  $\frac{\ud B}{\ud t} $ in the above model. The effects of benefits (higher assimilation) and costs (higher respiration) of increased Area occur within the same element of eq. \ref{eq:eq0}, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light.

@@

\subsubsection*{Stage effect}\label{stage-effect}

>>>>

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theoretical work, and test these predictions in tree species using a meta-analysis of 103 studies ($>$ 500 correlations) for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate). We also recorded data for 14 other traits commonly used in the trait literature. To capture the effects of plant size, we tested for a shift in the direction of correlation between growth rates and traits across three ontogenetic stages: seedling, sapling and adult.

\item Results were consistent with predictions, although there were some limitations arising from unequal numbers of observation across ontogenetic stages. Specific leaf area was correlated with relative growth rate in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by ontogenetic stage. Seed mass and maximum height were correlated with relative growth rate only in one ontogenetic stage category: seedlings and adults, respectively.

\item Although we were able to confirm several of our theoretical predictions, major knowledge gaps still exist in the trait literature. For example, for one-third of the traits considered, the majority ($>$ 75\%) of reported correlations with growth came from the same ontogenetic stage.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rate  among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. During the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. This approach, using measurable traits, made broad comparisons possible across continents, latitudes, and thousands of species. Since then, functional traits have been seen as constructive tools to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of trait variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, recently a number of empirical studies indicate the correlation between traits and growth is not consistent. Consider SLA as an example. By definition, species with high SLA can deploy more leaf area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In seedlings growing under favourable conditions, SLA has repeatedly been found to be a strong predictor of relative growth rate, regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{King-1999, Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Size has long been known to play a central role in organism design and function \citep{Vogel:1988ux, Vogel:2003wb}, including that of trees \citep{Farnsworth:1995im, Givnish:1995ta, King-2011}. As plants grow larger, a greater amount of energy is devoted towards building and maintaining non-photosynthetic tissues \citep{Givnish:1995ta,King-2011}. A recent model by \citet{Falster:2011ii} suggests this change shifts the benefits and costs of some trait-based trade-offs, such that the net effect of the trait on growth changes with size. Thus the absence of a correlation between SLA and growth at larger sizes may in fact be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii}. While recent empirical studies support the idea that the relationship between traits and growth rates may change with size (\citealt{Iida:2014ep, Iida:2014hq}), the trend and mechanisms behind this variation remain unclear.

In this paper, we consider five prominent traits (specific leaf area = SLA, wood (or stem) tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum (or asymptotic height), and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

We first outline equations for the new growth model, adapted from \citet{Falster:2011ii}. We then use this model to generate qualitative predictions for how the effect of traits on growth rate changes with plant size. These hypotheses are summarised in Table \ref{tab:trade-offs}.

@@

The rate at which \textbf{live biomass} $B$ is produced by a plant is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}} \, ,
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i = \rm{l,s,r,a,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes \textbf{areas}, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm a} + M_{\rm f}$.  Applying the chain rule of differential calculus, the \textbf{growth rate in plant mass}, ($M_{\rm a} $), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm a} }{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm a} }{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm a} }{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The \textbf{growth rate in height} can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm a} }}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}.
\end{equation}
where $\ud H / \ud A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area. This term expresses the architectural
strategy of the plant. We assume it is unrelated to the traits considered in the current paper, in other words that it does not automatically change in response to changes in the traits. The second term, $\ud A_{\rm l} / \ud M_{\rm a} $,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures.  $\ud A_{\rm l} / \ud M_{\rm a} $
can itself be expressed as a sum of construction costs per unit leaf
area,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm a} }
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

Similarly \textbf{growth rate in basal area} can be expressed as the sum of growth in sapwood and heartwood, with the former term again decomposed as a product:
\begin{equation}\label{eq:eq3}
\begin{array}{ccccc}
\underbrace{{\frac{\ud A_{\rm st}}{\ud t}}}_{\text{basal area growth rate}} &=& \frac{\ud A_{\rm ss}}{\ud t} &+& \frac{\ud A_{\rm sh}}{\ud t} \\
&=& \underbrace{\frac{\ud A_{\rm ss}}{\ud A_{\rm l}}  \times {\frac{\ud A_{\rm l}}{\ud M_{\rm a} }} \times  {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}}_{\text{increase in total sapwood}} &+&  \underbrace{k_{\rm ss} \times \frac{\ud A_{\rm ss}}{\ud t}}_{\text{sapwood turnover}},\\
\end{array}
\end{equation}
where $\frac{\ud A_{\rm ss}}{\ud A_{\rm l}} $ is the  sapwood area per leaf area. The \textbf{growth rate in diameter} is then obtained via the geometric relationship between stem diameter and stem area,
\begin{equation}\label{eq:eq4}
\underbrace{\frac{\ud D}{\ud t}}_{\text{growth rate in diameter}} = \sqrt[]{\frac{\Pi}{A_{\rm st}}} \times {\frac{\ud A_{\rm st}}{\ud t}}.
\end{equation}

@@

The above equations are general, i.e. they do not depend on detailed assumptions about plant construction. To make qualitative predictions with respect to traits such as specific leaf area and wood density we require one additional assumption: that the mass of sapwood tissue ($M_{\rm ss}$) increases disproportionately with respect to the plant's leaf area ($A_{\rm l}$). This assumption is included in nearly every model of plant growth that has been developed, on the basis that as stems grow longer the sapwood associated with each leaf must be increasing. It has the effect of changing the relative importance of different components within equations for $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area) as a plant increases in size. Specifically, in eq. \ref{eq:eq0}, the costs of stem respiration and turnover increase faster than leaf photosynthesis, respiration or turnover, as the plant grows. Similarly, in eq. \ref{eq:daldmt}, the costs of constructing sapwood to support a given unit of leaf area, $\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$, increases faster than the cost of producing a unit of leaf area, $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ which is simply the inverse of specific leaf area. Whether plant growth rate increases with a specific trait, therefore depends on how the trait affects the different elements of \ref{eq:eq0}-\ref{eq:eq4} and how those effects change with size (Table \ref{tab:trade-offs}).

Based on the above model and this single assumption about growth, we outline five hypotheses that arise by considering the influence of trait variation on elements of equations  \ref{eq:eq0}-\ref{eq:eq4} and how these may interact with plant size (see Table \ref{tab:trade-offs}). It is possible that other correlations may arise indirectly. Nevertheless, for the purposes of this paper we adopt as working hypotheses only those arising from direct mechanistic causation. We consider predictions for each of height, diameter, and mass-based growth, and how they  may vary depending whether growth is expressed in absolute or relative terms, (see Table \ref{tab:trade-offs}). For many trait and growth combinations, the predicted effect of the trait on growth is similar when expressed in relative and absolute terms. However, this is not true for all traits (details given below).
\subsubsection*{H1: The relationship between SLA and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants}

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area), appearing on the right side in the equations for growth rate (\ref{eq:eq1}, \ref{eq:eq2}, \ref{eq:eq03}). Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA changes with size (Table \ref{tab:trade-offs}).
During early life, when leaves comprise a large part of the plant, increasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ (the cost of producing a unit of leaf area) has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is negative. This occurs because of the increased turnover of high SLA species.

@@

Assuming plants otherwise maintain a similar cross section of conducting tissue, cheaper stem construction (low wood density or stem density) will increase the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$), and consequently growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass when expressed in relative terms. However, when expressed as absolute growth, the predicted relationship is non-significant along all plant size \citep[see also][]{Enquist-1999}.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm a} }{\ud B}$ (the allocation to vegetative growth) in eqs. (\ref{eq:eq2}-\ref{eq:3}). For seedlings, where all individuals are investing solely in growth, we expect no influence of maximum height on growth (i.e. $\frac{\ud M_{\rm a} }{\ud B} =1$). But once some species begin to produce seed, then species closer to their maximum height value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller maximum height are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of growth is used (Table \ref{tab:trade-offs}).

@@

The only effect of increased seed size in the above model is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have faster absolute growth rates. However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based growth rates during the earliest phases of growth \citep{Turnbull:2012ew}. Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therefore disappear. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

\subsection*{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, particularly under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is typically measured. This effect occurs by increasing  $\frac{\ud B}{\ud t} $ in eqs. (\ref{eq:eq1}-\ref{eq:eq3}). The positive effect of increasing Aarea (higher assimilation) may be offset by the costs (higher respiration), however, assuming both increase in proportion, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light.

@@

\subsubsection*{Ontogenetic stage effect}\label{stage-effect}
